### 2023-06-26 paper/paper.tex
chore: Initial JCon draft

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

\verb|AdaptiveHierarchicalRegularBinning.jl| computes a hierarchical space-partitioning
tree for a given set of points of arbitrary dimensions, that divides the space and
stores the reordered points offering efficient access. Space-partitioning data
structures are vital for algorithms that exploit spatial distance to reduce
computational complexity, see for example the Fast Multipole Method, and algorithms
finding nearest neighbors and their applications.

\end{abstract}

\section{Overview}
To partition a cloud of points, \verb|AdaptiveHierarchicalRegularBinning.jl| maps each
point to a $1$ dimensional space filling curve, Morton or Z-order curve. The hierarchical
tree is defined by the Morton index of each point.

\begin{example}
  For a $2$-dimensional space a target spatial tree with $2$ levels (other than the root).
  Assume we compute a morton index of $(0100)_2$. The 2 most significant bits address the
  top most level of the tree (the one directly under root).

  The most significant bit
  addresses the first dimension and the value $(0)_2$ signifies that the point lies in
  the left-most half of the dimension.

  The second most significant bit addresses the
  second dimension and a value of $(1)_2$ signifies that the point lies in the right-most
  half of the dimension (typicaly the lower-most half since this is $2$-dimensional,
  right-most is a more generic term used for any number of dimensions).

  % TODO Insert gaphs
\end{example}

Sorting in terms of these morton indices group the points togather for efficient access
when the tree is iterated over. One natural way of sorting these indices is the
\verb|Radixsort| where we sort based on the most significant radix and recurse for the
remaining bits in the same fashion. This type of sorting process respects the nature of
the morton index and thus can offer benefits over other sorting algorithms.

One benefit of the \verb|Radixsort| is that it offers partial sorting capabilities. When
the number of items within a bin is not high enough, there is no reason to continue the
spatial partitioning and therefore there is no reason to continue sorting. Stopping
\verb|Radixsort| at a specific level of recursion translates to stoping the tree at that
level, there is a $1-1$ corespondance of the \verb|Radixsort| recursion and the tree.

\verb|Radixsort| can also be trivialy parallelized, after sorting the based on the most
significant radix the sub-arrays left to sort are mutually independant and thus offer
themselves to easy parallelization.

Sorting based on a radix is not trivial. Any sorting algorithm will work just fine but
since small radix means small number of discrete values \verb|Countsort| is used.
\verb|Countsort| offers trades-off spatial complexity for runtime complexity.
The worst case runtime complexity of the algorithm is $O(n)$, which means that
it scales linearly with the size of the array to sort. While the spatial complexity for
the in-place variant is $O(max(r))$ where $r$ is the selected radix and $max(r)$ is the
maximum value allowed by $r$.

\verb|Countsort| can be parallelized as well, though not trivialy. The parallel variant
requires a larger spatial complexity of $O(t\cdot max(r))$ where $t$ is the number of
tasks allocated for the parallelization. The runtime complexity drops by the same factor
$t$, $O(n/t)$. Spatial complexity is a limites the capabilities of the algorithm but even
for large values of $t$ this auxilary space needed is not a bad trade-off.

An other trade-off to note is the selection of radix. The bigger the radix the larger
the load on the \verb|Countsort|. Since the time complexity of \verb|Countsort| is $O(n)$
and even better for the parallel variant $O(n/t)$ we prefer that most of the work is
delegated to \verb|Countsort| instead of the recursion process of \verb|Radixsort|.
Selecting a large radix though, might have an adverse effect. The auxilary space needed
is based on the max value of the radix which exponatialy grows as the radix grows, meaning
that we quickly run out of memory for larger matrices.

Our implementation is radix-agnostic, meaning that the user can select the radix. But the
default approach is a radix of $16$ bits. $16$ bits means that at the worst case of
morton indices of $128$ bits we will need to recurse $8$ times and the auxilary space needed
is a factor of $2^{16} = 64kB$.

\section{Usage}

\begin{lstlisting}[language = Bash]
  $ julia -t auto
\end{lstlisting}

\begin{lstlisting}[language = Julia]
  using AdaptiveHierarchicalRegularBinning
  n = 10000
  d = 5
  X = rand(n, d)

  tree = regular_bin(X, dpt, sml_th; dims=2)
  # X is no permuted to accomodate morton encoding

  for leaf in Leaves(tree)
  # iterates over the tree's leafs
  end

  for node in PreOrderDFS(tree)
    # iterates over the tree's nodes in pre order
  end

  I, D = knn(tree, X)

  original_perm!(tree, I, D)
  # I and D are now refering to the
  # original permutation
\end{lstlisting}

\section{Theory}
Assuming a set of $n$ $d$-dimensional points $\mathcal{V}$, we represent $\mathcal{V}$ as
the $d \times n$ matrix $V$. \verb|AdaptiveHierarchicalRegularBinning| uses Morton
encoding to map each $d$-dimenensional point of $\mathcal{V}$ to a $1$-dimenensional
space filling curve, Z-order curve.

The mapping proccess requires 3 distinct steps.

\begin{itemize}
  \item Normalize $\mathcal{V}$ such that each vector $\vec{v}$ of $\mathcal{V}$
  satisfies $\vec{v} \in [0,1)^d$
  \item Quantize each dimension using an $l$-bit quantizer
  \item Bit-interleave the output of the quantizer
\end{itemize}

\subsection{Normalization}
In order to compute the hierarchical tree we first normalize $\mathcal{V}$.

$$ \vec{v} \in \mathcal{H}, \forall \vec{v} \in \mathcal{V}$$

Where $\mathcal{H}$ is the connected $d$-dimensional space that contains all vectors
of $\mathcal{V}$.

$$ H \triangleq [m_1, M_1] \times [m_2, M_2] \times \cdots \times [m_d, M_d] $$

Where
$$\space m_i = \min_j\left(V_{i,j}\right), \forall i \in \left\{1, 2, \cdots, d\right\}$$
$$\space M_i = \max_j\left(V_{i,j}\right), \forall i \in \left\{1, 2, \cdots, d\right\}$$

Therefore the goal of the normalization process is to define a $1-1$ function $\hat{f}$
such that

$$\hat{f}: \mathcal{H} \rightarrow [0, 1)^d$$

Applying the function to each element of $\mathcal{V}$ defines a new set of points
$\hat{\mathcal{V}}$ the elements of which are by definition inside a $d$-dimensional unit
hypercube.

$$ \hat{\mathcal{V}} \triangleq \left\{ \hat{f}\left(v\right) : v \in \mathcal{V} \right\}  $$
$$ \vec{v} \in [0, 1)^d, \forall \vec{v} \in \hat{\mathcal{V}} $$

Furthermore $\hat{f}$ must satisfy one more property. Namely, relative distances
must be preserved.

$$ ||\vec{v_1} - \vec{v_2}|| \geq ||\vec{v_1} - \vec{v_3}|| \Leftrightarrow ||\hat{f}(\vec{v_1}) - \hat{f}(\vec{v_2})|| \geq ||\hat{f}(\vec{v_1}) - \hat{f}(\vec{v_3})||$$
$$ \forall \vec{v_1}, \vec{v_2}, \vec{v_3} \in \mathcal{V} $$

Given this constraint we can therefore define the function $\hat{f}$ as such
\begin{definition}
  $$\hat{f}(\vec{v}) \triangleq \frac{\vec{v} - \vec{t}}{s}, \forall \vec{v} \in \mathcal{H} \quad \textrm{where}\; \vec{t} \in \mathbb{R}^d, s\in\mathbb{R}^*$$
\end{definition}

It can be proven that this definition satisfies the constraint

\begin{lemma}
  $$
  \begin{matrix}
    ||\hat{f}(\vec{v_1}) - \hat{f}(\vec{v_2})|| &=& || \frac{\vec{v_1} - \vec{t}}{s} - \frac{\vec{v_2} - \vec{t}}{s}||\\
                                                &=& || \frac{\vec{v_1} - \vec{t} - \vec{v_2} + \vec{t}}{s}||\\
                                                &=& || \frac{\vec{v_1} - \vec{v_2}}{s}||\\
                                                &\triangleq& \frac{||\vec{v_1} - \vec{v_2}||}{|s|} & \forall v_1, v_2 \in \mathcal{V}\\
  \end{matrix}
  $$
\end{lemma}
\begin{proof}
  $$
  \begin{matrix}
    ||\hat{f}(\vec{v_1}) - \hat{f}(\vec{v_2})|| \geq ||\hat{f}(\vec{v_1}) - \hat{f}(\vec{v_3})|| &\Leftrightarrow&\\
    \frac{||\vec{v_1} - \vec{v_2}||}{|s|} \geq \frac{||\vec{v_1} - \vec{v_3}||}{|s|} &\Leftrightarrow&\\
    ||\vec{v_1} - \vec{v_2}|| \geq ||\vec{v_1} - \vec{v_3}||
  \end{matrix}
  $$
$$$$
\end{proof}

The effect that $\vec{t}$ and $s$ have on the space $\mathcal{H}$ can be computed as such
$$
\begin{matrix}
  \hat{f}(\mathcal{H}) & = & \bigtimes\limits_{i=1}^d [\hat{f}(m_i), \hat{f}(M_i)] &\\
                       & = & \bigtimes\limits_{i=1}^d [\frac{m_i-\vec{t}_i}{s}, \frac{M_i-\vec{t}_i}{s}] &\
\end{matrix}
$$

Defining $\vec{t}$ and $s$ is trivial.

$$ \vec{t}_i = \begin{pmatrix} m_1 \\ m_2 \\ \vdots \\ m_d \end{pmatrix}$$
$$ s = \max_i(M_i-m_i)$$

Therefore $\hat{f}(\mathcal{H}) \subseteq [0, 1]^d \Leftrightarrow \hat{f}: \mathcal{H} \rightarrow [0,1]^d$

A subtle, yet important, modification is made to the computed $s$ in order to truely
map each dimension to $[0,1)$ instead of $[0,1]$.

$$ s' = s + \epsilon(s) $$

We define $\epsilon(s)$ as the smallest possible value such that $s + \epsilon(s) > s$.
In mathematics when $s \in \mathbb{R}$ there is no such value, due to the uncountably
infinite nature of $\mathbb{R}$. But computers use finite bits to represent their
floating point values and so $\epsilon(s)$ is the machine's epsilon value.

\subsection{Quantization}
We use an $l$-bit quiantizer to quantize $\hat{\mathcal{V}}$.
$$q^{(l)}(x) = \lfloor x \cdot 2^l \rfloor, \forall x \in [0, 1)$$

This trivialy generalizes to $d$ dimensions.
$$\vec{q^{(l)}}(\vec{x}) = \begin{pmatrix}
  q^{(l)}(x_1)\\
  q^{(l)}(x_2)\\
  \vdots \\
  q^{(l)}(x_d)
\end{pmatrix}, \forall \vec{x} \in [0,1)^d$$

Therefore each dimension is mapped to an $l$-bit word. The bits of
this word approximate the location of the point in the given
dimension.

The output of the quantizer can be mapped $1-1$ to a value in the
original space $[0,1]$. This real value representation can be
trivialy computed as such
$$ \tilde{x}_l = \frac{q^{(l)}(x)}{2^l} $$
Since $2^l-1$ is the max value of the quantizer.

Assume that $q^{(l)}(x) = (b_1b_2b_3\dots b_l)_2$ where $b_1$ is the most significant
bit of the output word and $b_l$ the least significant.
$$
\begin{matrix}
  \tilde{x}_l & = & \frac{q^{(l)}(x)}{2^l} \\
              & = & \frac{(b_1b_2b_3\dots b_l)_2}{2^l} \\
              & = & \frac{\sum\limits_{i=1}^{l} b_i \cdot 2^{l-i}}{2^l} \\
              & = & \sum\limits_{i=1}^{l}\frac{b_i \cdot 2^{l-i}}{2^l} \\
              & = & \sum\limits_{i=1}^{l}b_i \cdot 2^{-i} \\
\end{matrix}
$$

So then, each bit controls a diminishing power of $2$. The geometric interpretation of
this is that
\begin{itemize}
  \item The most significant bit indicates in which half the real value lies, $1$
  for the right-most half and $0$ for the left-most.
  \item The second most significant bit indicates the same but for the previously
  selected half's half.
  \item And so on.
\end{itemize}

\subsection{Morton encoding}
Computing the Z-order curve from the output of the quantizer is done via bit-interleaving.
Assuming that for some input $\vec{x}$
$$
\vec{q}^{(l)}(\vec{x}) =
\begin{pmatrix}
  (b_{11}b_{12}b_{13}\dots b_{1l})_2\\
  (b_{21}b_{22}b_{23}\dots b_{2l})_2\\
  \vdots \\
  (b_{d1}b_{d2}b_{d3}\dots b_{dl})_2
\end{pmatrix}
$$

Interleaving the bits would compute the following binary value.

$$(b_{11}b_{21}\dots b_{d1}\quad b_{12}b_{22}\dots b_{d2}\quad \dots \quad b_{1l}b_{2l}\dots b_{dl})_2$$

This binary value is semanticaly divided into $l$ groups of $d$ bits. Each bit in a group
describe the location of the input vector $\vec{x}$ in the respective dimension, the
first bit describes the first dimension, the second bit the second dimension and so on.
Each group of bits describe wether the value is located in the left-most or the
right-most part of the respective level, the first group describe the whole space
$[0,1)^d$, the second group describe the selected half of the first group and so on.

\section{Limitations}

\subsection{Curse of high dimensionality}
Data of high dimensionality exponatialy get sparser as the number of dimensions grow.
The sparser the data the lower the probabilities are for two points to occupy the same
bin. In actuallity after the first recursion of the algorithm each element is most
probably assigned a unique bin, which defeats the purpose of the subdivision.

\subsection{Morton word size}
To compute the Z-order curve we first quantize the real values with an $l$-bit
quantizer and then we interleave $d$ words of $l$ bits. The resulting binary value has in
total $l\cdot d$ bits which are stored as an \verb|Unsigned| value, limiting the size of the
value to $128$ bits, which is the biggest \verb|Unsigned| built-in integer.

\subsection{Space complexity}
We perform the partial sorting of the $V$, the cloud of points in an out-of-order fashion,
which marks the space complexity needed for the algorithm as $O(n\cdot d)$ where $n$ is
the number of points and $d$ the number of dimensions. An in-place implementation would
trade-off time complexity for space complexity.

\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2023-07-01 paper/paper.tex
chore: Overview

@@
 \section{Overview}
-To partition a cloud of points, \verb|AdaptiveHierarchicalRegularBinning.jl| maps each
-point to a $1$ dimensional space filling curve, Morton or Z-order curve. The hierarchical
-tree is defined by the Morton index of each point.
+To partition a cloud of points of $d$ dimensions,
+\verb|AdaptiveHierarchicalRegularBinning.jl| maps each point to a $1$ dimensional
+space filling curve, Morton or Z-order curve. The hierarchical tree is defined by the
+Morton index of each point.
 
-\begin{example}
-  For a $2$-dimensional space a target spatial tree with $2$ levels (other than the root).
-  Assume we compute a morton index of $(0100)_2$. The 2 most significant bits address the
-  top most level of the tree (the one directly under root).
+\subsection{Morton encoding}
+Morton indices are binary words grouped into $l$ levels of $d$ bits each. Each level
+describes in an increasingly finer detail the area in which the point resides. The
+first level, comprised of the $d$ most significant bits of the morton index, describes
+only in which half of each dimension the point lies.
 
-  The most significant bit
-  addresses the first dimension and the value $(0)_2$ signifies that the point lies in
-  the left-most half of the dimension.
+$$b = \left\{\begin{matrix}
+  (0)_2, & \textrm{Left-most half}  \\
+  (1)_2, & \textrm{Right-most half}
+\end{matrix}\right.$$
 
-  The second most significant bit addresses the
-  second dimension and a value of $(1)_2$ signifies that the point lies in the right-most
-  half of the dimension (typicaly the lower-most half since this is $2$-dimensional,
-  right-most is a more generic term used for any number of dimensions).
+\begin{definition}[Left/Right-mostness]
+  For a $1$-dimensional simply connected space $A = [m, M)$, we say that
+  $x$ lies in the Left-most half, if and only if, $x<h=\frac{M+m}{2}$.
+  Otherwise $x$ lies in the Right-most half.
+
+  $$
+  \begin{matrix}
+    L &\triangleq& [m, h) \\
+    R &\triangleq& [h, M)
+  \end{matrix}
+  $$
+
+  \begin{figure}
+    \centerline{
+      \includegraphics[width=10pc]{figures/spatial-division.png}
+      \includegraphics[width=10pc]{figures/spatial-tree.png}
+    }
+    \caption{ (a) Spatial division, (b) Spatial tree }
+    \label{spatial-tree}
+  \end{figure}
+
+\end{definition}
+
+\begin{example}[Morton index interpretation]
+  Figure \ref{spatial-tree} depicts a $2$-dimensional space and a target spatial
+  tree of at most $2$ levels (other than the root). The morton index, in such a tree,
+  will have a length of $2\cdot2=4$. Assume we compute a morton index of $(0011)_2$.
+  The first group is comprised of the $2$ most significant bits and addresses the first
+  level of the tree (the one directly under root).
+  \\\\
+  The most significant bit addresses the first dimension and the second most significant
+  bit addresses the second dimension. The value $(0)_2$ signifies that the point lies in
+  the left-most half of each dimension.
+  \\\\
+  The next group describes the same quantities only a level lower. Left/Right-mostness
+  is defined in the context of that level. Both values are $(1)_2$ so the point lies in
+  the right-most half of both dimensions in the second level.
+  \\\\
+  Summarizing, the point belongs to the:
+  \begin{itemize}
+    \item right-most half of the left-most half of the first dimension.
+    \item right-most half of the left-most half of the second dimension.
+  \end{itemize}
 
@@
 
-Sorting in terms of these morton indices group the points togather for efficient access
-when the tree is iterated over. One natural way of sorting these indices is the
-\verb|Radixsort| where we sort based on the most significant radix and recurse for the
-remaining bits in the same fashion. This type of sorting process respects the nature of
-the morton index and thus can offer benefits over other sorting algorithms.
+\subsection{Reordering}
+Sorting in terms of the morton indices groups the points together for efficient access
+when the tree is iterated over. We use the \verb|MSD Radixsort| algorithm to perform the
+soring.
 
-One benefit of the \verb|Radixsort| is that it offers partial sorting capabilities. When
-the number of items within a bin is not high enough, there is no reason to continue the
-spatial partitioning and therefore there is no reason to continue sorting. Stopping
-\verb|Radixsort| at a specific level of recursion translates to stoping the tree at that
-level, there is a $1-1$ corespondance of the \verb|Radixsort| recursion and the tree.
+\subsubsection{MSD Radixsort}
+\verb|MSD Radixsort| sorts based on the most significant radix. Afterwards, elements
+with the same most significant digit (MSD) form a consecutive sub-array. The algorithm
+recurses for each such sub-array selecting the next digit in line. Selecting a large
+radix is crucial for the algorithm since a large radix lowers the number of recursions.
+\\\\
+\verb|MSD Radixsort| respects the hierarchical nature of the target tree and thus it offers
+benefits over other sorting algorithms. Each recursion of the algorithm sorts a section
+of the resulting tree. If the radix matches the number of dimensions then each recursion
+sorts the current level of the tree.
+\\\\
+When the number of elements of a sub array is small the recursion for that sub-array stops.
+This leads to a partialy sorted array by design. Not sorting small sub-arrays leads to less
+overhead without sacrificing future performance. Nodes with unsorted morton indices will
+always be leaf nodes.
+\\\\
+\verb|MSD Radixsort| can be trivialy parallelized. After the initial sorting, the formed
+sub-arrays are mutually exclusive, meaning that no data dependencies are shared between
+them. Each sub-array can then be sorted in parallel.
 
-\verb|Radixsort| can also be trivialy parallelized, after sorting the based on the most
-significant radix the sub-arrays left to sort are mutually independant and thus offer
-themselves to easy parallelization.
+\subsubsection{Countsort}
+\verb|MSD Radixsort| needs a sorting algorithm to sort the selected radix in each recursion.
+\verb|Countsort| is a great fit for this type of sorting. \verb|Countsort| trades
+space complexity for runtime complexity. The algorithm uses an auxilary vector, whose
+size scales linearly with the selected radix $O(r)$. The benefit is the runtime
+complexity is order $O(n + r)$ where $n$ is the size of the array to sort. Typically, $n \gg r$
+and so the runtime complexity reduces to $O(n)$.
+\\\\
+\verb|Countsort| can be parallelized as well, though not as trivialy. The parallel version
+of \verb|Countsort| requires space complexity of $O(t\cdot r)$ and runtime complexity of
+$O(n/t + t\cdot r)$ where $t$ is the number of threads in the system. As long as $n$ is
+sufficiently large $n \gg t^2\cdot r$ then the parallel implementation is preferable since
+the runtime complexity reduces to $O(n/t)$.
 
-Sorting based on a radix is not trivial. Any sorting algorithm will work just fine but
-since small radix means small number of discrete values \verb|Countsort| is used.
-\verb|Countsort| offers trades-off spatial complexity for runtime complexity.
-The worst case runtime complexity of the algorithm is $O(n)$, which means that
-it scales linearly with the size of the array to sort. While the spatial complexity for
-the in-place variant is $O(max(r))$ where $r$ is the selected radix and $max(r)$ is the
-maximum value allowed by $r$.
+\subsubsection{Radix trade-off}
+While \verb|MSD Radixsort| fairs better for larger radices, \verb|Countsort| reduces in
+performance.
 
-\verb|Countsort| can be parallelized as well, though not trivialy. The parallel variant
-requires a larger spatial complexity of $O(t\cdot max(r))$ where $t$ is the number of
-tasks allocated for the parallelization. The runtime complexity drops by the same factor
-$t$, $O(n/t)$. Spatial complexity is a limites the capabilities of the algorithm but even
-for large values of $t$ this auxilary space needed is not a bad trade-off.
-
-An other trade-off to note is the selection of radix. The bigger the radix the larger
-the load on the \verb|Countsort|. Since the time complexity of \verb|Countsort| is $O(n)$
-and even better for the parallel variant $O(n/t)$ we prefer that most of the work is
-delegated to \verb|Countsort| instead of the recursion process of \verb|Radixsort|.
-Selecting a large radix though, might have an adverse effect. The auxilary space needed
-is based on the max value of the radix which exponatialy grows as the radix grows, meaning
-that we quickly run out of memory for larger matrices.
-
-Our implementation is radix-agnostic, meaning that the user can select the radix. But the
-default approach is a radix of $16$ bits. $16$ bits means that at the worst case of
-morton indices of $128$ bits we will need to recurse $8$ times and the auxilary space needed
-is a factor of $2^{16} = 64kB$.
+\begin{example}
+  For a $128$-bit morton index, a radix of $16$-bits is selected.
+  The number of recursions in this case is $8$ which is fairly small and the condition
+  $n\gg t^2\cdot r$ can still be satisfied.
+  % TODO: Graphs runtime complexity vs r
+\end{example}
 
@@
 trade-off time complexity for space complexity.
+
+\section{Experiments}
+
+\subsection{Uniform data 2 dimensions}
+\subsection{Uniform data 10 dimensions}
+\subsection{MNIST data 10 dimensions}
+\subsection{PBMC data 10 dimensions}
+\subsection{PBMC 36 dimensions}
 

### 2023-07-01 paper/paper.tex
chore: Additions to paper

@@
   $$
-
-  \begin{figure}
-    \centerline{
-      \includegraphics[width=10pc]{figures/spatial-division.png}
-      \includegraphics[width=10pc]{figures/spatial-tree.png}
-    }
-    \caption{ (a) Spatial division, (b) Spatial tree }
-    \label{spatial-tree}
-  \end{figure}
-
 \end{definition}
@@
 
-  % TODO Insert gaphs
+  \begin{figure}
+    \centerline{
+      \includegraphics[width=10pc]{figures/spatial-division.png}
+      \includegraphics[width=10pc]{figures/spatial-tree.png}
+    }
+    \caption{ (a) Spatial division, (b) Spatial tree }
+    \label{spatial-tree}
+  \end{figure}
 \end{example}
@@
 Data of high dimensionality exponatialy get sparser as the number of dimensions grow.
-The sparser the data the lower the probabilities are for two points to occupy the same
-bin. In actuallity after the first recursion of the algorithm each element is most
-probably assigned a unique bin, which defeats the purpose of the subdivision.
+The sparser the data the lower the probabilities are for any two points occupying the same
+bin.
+\\\\
+Assuming uniformly distributed $d$-dimensional $n$ points and a target tree of
+$l$ levels, the expected number of points per leaf node is $E[n_\textrm{leaf}]$.
+
+$$E[n_\textrm{leaf}] = \frac{n}{2^{l\cdot d}}$$
+
+The expected number of points per leaf node drops exponentialy with $d$. This extends
+to most distributions of points since the probability of a point belonging to a node is a
+function of the $d$-dimensional area that the node occupies $A^{(d)}_l $.
+In the case of uniform distribution the probability is directly proportional
+to this area.
+
+$$A^{(d)}_l = 2^{-l \cdot d}$$
+
+In general the area is $A^{(d)}_l = s\cdot 2^{-l \cdot d}$ where $s$ is the scaling factor
+of the original space. But since scaling does not affect the relative density of the
+points the probability is not affected.
 
@@
 \section{Experiments}
+\subsection{Dimensionality scaling}
+For $n=10^5$ uniformly distributed points we computed the average number of points in a
+leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
+dimensional spaces the data get exponatialy sparser.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=20pc]{figures/experiments/b_vs_d.png}
+  }
+  \caption{ Exponential decrease of mean number of points per node }
+  \label{b_vs_d}
+\end{figure}
 
 \subsection{Uniform data 2 dimensions}
+
+\begin{lstlisting}[language = Julia]
+julia> d, n = size(X)
+(2, 100000)
+
+julia> l = 4;
+
+julia> @time t = regural_bin(UInt128,
+                             X,
+                             l,
+                             1;
+                             dims=2);
+0.025399 seconds (... 43.93% gc time)
+\end{lstlisting}
+
+For uniformly distributed data we expect that each leaf node will contain
+$n_{\textrm{node}}$ points.
+$$n_{\textrm{node}} = \frac{n}{2^{l\cdot d}}$$
+
+In this case $n = 10^5$, $d = 2$ and $l = 4$. So the expected number of points within a
+node is $n_{\textrm{node}} = \frac{10^5}{2^{4\cdot 2}} =390.625$.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=20pc]{figures/uniform_2d/hist.png}
+  }
+  \caption{ Distribution of node length }
+  \label{uniform-2d-hist}
+\end{figure}
+
+The curse of high dimensionality can be deduced by the formula of $n_{\textrm{node}}$.
+The number of possible leaf nodes is $2^{l\cdot d}$ which grows exponentialy with $l$ and
+$d$, meaning that for larger $d$ the total number of points $n$ must also grow exponentialy
+to keep the ratio constant.
+
 \subsection{Uniform data 10 dimensions}
+
+\begin{lstlisting}[language = Julia]
+julia> d, n = size(X)
+(10, 100000)
+
+julia> l = 1;
+
+julia> @time t = regural_bin(UInt128,
+                             X,
+                             l,
+                             1;
+                             dims=2);
+0.046737 seconds (2.48 k allocations: 30.789 MiB)
+\end{lstlisting}
+
 \subsection{MNIST data 10 dimensions}

### 2023-07-02 paper/paper.tex
chore: Thread scaling additions to paper

@@
 \end{example}
+
+Our implementation utilizes a watefall scheme to adapt to the needs of the problem at
+hand. There are $3$ different thresholds that can tune the runtime performance of the
+spatial partitioning.
+
+\begin{itemize}
+  \item \verb|Huge| threshold
+  \item \verb|Big| threshold
+  \item \verb|Small| threshold
+\end{itemize}
+
+$$ \textrm{Huge} > \textrm{Big} > \textrm{Small} $$
+
+For sub-arrays that are larger than the \verb|Huge| threshold our implementation
+uses parallel \verb|MSD Radixsort|, otherwise it falls back to the non parallel variant.
+\\\\
+For sub-arrays that are larger than the \verb|Big| threshold our implementation uses
+parallel \verb|Countsort|, otherwise it falls back to the non parallel variant.
+\\\\
+For sub-arrays that are smaller than the \verb|Small| threshold our implementation
+stops the recursion of \verb|MSD Radixsort|.
 
@@
 \section{Experiments}
+All conducted experiments are executed with a selected radix of $16$ bits and the array
+is fully sorted.
+
 \subsection{Dimensionality scaling}
@@
   \label{b_vs_d}
+\end{figure}
+
+\subsection{Runtime performance}
+\subsubsection{Scaling with number of points}
+For $l=4$ max tree depth and $d=4$ number of dimensions
+\verb|AdaptiveHierarchicalRegularBinning.jl| scales linearly with the size of the array
+$n$.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=20pc]{figures/experiments/scale_n.png}
+  }
+  \caption{ Linear runtime scaling }
+  \label{scale_n}
+\end{figure}
+
+\subsubsection{Runtime dimensionality scaling}
+High dimensionality should affect the morton encoding since bit-interleaving scales
+linearly with the dimensions. The reordering should be affected slightly since each
+element
+
+% TODO
+
+\subsubsection{Scaling with number of threads}
+For $n=10^6$ points of $d=10$ dimensions and a target tree of $l=4$
+levels, we can see from \ref{scale_t} that the scaling of
+\verb|AdaptiveHierarchicalRegularBinning.jl| drops exponatialy
+with the number of threads.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=20pc]{figures/experiments/scale_t.png}
+  }
+  \caption{ Parallelization scaling }
+  \label{scale_t}
 \end{figure}

### 2023-07-02 paper/paper.tex
chore: Additions to paper

@@
 
+\subsection{Tree}
+The resulting tree is memory-dense, meaning that it does not store empty nodes.
+The root node encapsulates the entire cloud of points, while each child node points to
+a consecutive sub-array of those points. This characteristic holds true for every node
+within the tree. Each node is also aware if its level, parent and children.
+
 \section{Usage}
@@
 Assuming a set of $n$ $d$-dimensional points $\mathcal{V}$, we represent $\mathcal{V}$ as
-the $d \times n$ matrix $V$. \verb|AdaptiveHierarchicalRegularBinning| uses Morton
+the $d \times n$ matrix $V$. \verb|AdaptiveHierarchicalRegularBinning.jl| uses Morton
 encoding to map each $d$-dimenensional point of $\mathcal{V}$ to a $1$-dimenensional
 space filling curve, Z-order curve.
-
+\\\\
 The mapping proccess requires 3 distinct steps.
@@
   \item Normalize $\mathcal{V}$ such that each vector $\vec{v}$ of $\mathcal{V}$
-  satisfies $\vec{v} \in [0,1)^d$
-  \item Quantize each dimension using an $l$-bit quantizer
-  \item Bit-interleave the output of the quantizer
+  satisfies $\vec{v} \in [0,1)^d$.
+  \item Quantize each dimension using an $l$-bit quantizer, where $l$ is the maximum
+  depth of the target tree.
+  \item Bit-interleave the output of the quantizer.
 \end{itemize}
@@
 
-Furthermore $\hat{f}$ must satisfy one more property. Namely, relative distances
-must be preserved.
+Furthermore, the function $\hat{f}$ must satisfy an additional property, namely,
+the preservation of relative distances.
 
@@
 
-The effect that $\vec{t}$ and $s$ have on the space $\mathcal{H}$ can be computed as such
+We require the image of $\hat{f}$ to be $\hat{f}(\mathcal{H}) = [0,1)^d$. Thus we can
+compute $\vec{t}$ and $s$ to satisfy that constraint.
+
 $$
 \begin{matrix}
-  \hat{f}(\mathcal{H}) & = & \bigtimes\limits_{i=1}^d [\hat{f}(m_i), \hat{f}(M_i)] &\\
+  \hat{f}(\mathcal{H}) & \triangleq & \bigtimes\limits_{i=1}^d [\hat{f}(m_i), \hat{f}(M_i)] &\\
                        & = & \bigtimes\limits_{i=1}^d [\frac{m_i-\vec{t}_i}{s}, \frac{M_i-\vec{t}_i}{s}] &\
@@
 
-Defining $\vec{t}$ and $s$ is trivial.
+Hence,
 
@@
 Therefore $\hat{f}(\mathcal{H}) \subseteq [0, 1]^d \Leftrightarrow \hat{f}: \mathcal{H} \rightarrow [0,1]^d$
-
-A subtle, yet important, modification is made to the computed $s$ in order to truely
-map each dimension to $[0,1)$ instead of $[0,1]$.
+\\\\
+A subtle, yet crucial modification is made to the computed $s$ in order to
+accurately map each dimension to the interval $[0,1)$, rather than the inclusive range
+$[0,1]$.
 
@@
 We define $\epsilon(s)$ as the smallest possible value such that $s + \epsilon(s) > s$.
-In mathematics when $s \in \mathbb{R}$ there is no such value, due to the uncountably
-infinite nature of $\mathbb{R}$. But computers use finite bits to represent their
-floating point values and so $\epsilon(s)$ is the machine's epsilon value.
+In the standard real number system, when $s \in \mathbb{R}$, there is no such value,
+due to the uncountably infinite nature of $\mathbb{R}$. But computers use finite bits
+to represent floating point values and so $\epsilon(s)$ is the machine's epsilon value.
 
@@
 dimension.
-
+\\\\
 The output of the quantizer can be mapped $1-1$ to a value in the
-original space $[0,1]$. This real value representation can be
-trivialy computed as such
+original space $[0,1)$.
+
 $$ \tilde{x}_l = \frac{q^{(l)}(x)}{2^l} $$
-Since $2^l-1$ is the max value of the quantizer.
 
@@
 All conducted experiments are executed with a selected radix of $16$ bits and the array
-is fully sorted.
-
-\subsection{Dimensionality scaling}
+is fully sorted, unless stated otherwise.
+
+\subsection{Curse of high dimensionality}
 For $n=10^5$ uniformly distributed points we computed the average number of points in a
@@
   }
-  \caption{ Exponential decrease of mean number of points per node }
+  \caption{ Exponential decrease of mean number of points per leaf node }
   \label{b_vs_d}
@@
 \subsubsection{Scaling with number of points}
-For $l=4$ max tree depth and $d=4$ number of dimensions
-\verb|AdaptiveHierarchicalRegularBinning.jl| scales linearly with the size of the array
-$n$.
+For uniformly distributed $4$-dimensional data and $l=4$ max tree depth, the graph
+\ref{scale_n} suggests that \verb|AdaptiveHierarchicalRegularBinning.jl| scales
+linearly with the size $n$ of the array.
 
@@
 
-\subsubsection{Runtime dimensionality scaling}
-High dimensionality should affect the morton encoding since bit-interleaving scales
-linearly with the dimensions. The reordering should be affected slightly since each
-element
-
 % TODO
+% \subsubsection{Scaling with the number of dimensions}
+% High dimensionality should affect the morton encoding since bit-interleaving scales
+% linearly with the dimensions. The reordering should be affected slightly since each
+% element
 
 \subsubsection{Scaling with number of threads}
-For $n=10^6$ points of $d=10$ dimensions and a target tree of $l=4$
-levels, we can see from \ref{scale_t} that the scaling of
+For uniformly distributed $n=10^6$ points of $d=10$ dimensions and a target tree of
+$l=4$ levels, graph \ref{scale_t} suggests that the runtime of
 \verb|AdaptiveHierarchicalRegularBinning.jl| drops exponatialy
@@
 
+We would expect that for a large enough number of threads the performance would worsen,
+since the runtime complexity of the \verb|Countsort| algorithm suggests that if not
+$n \gg t^2 \cdot r$ then the parallel variant is worse than the sequential one. Our
+implementation gets around that limitation using the waterfall scheme for its
+parallelization and thus for small sub-arrays the sequential variant for the
+\verb|Countsort| is used.
+
 \subsection{Uniform data 2 dimensions}
@@
 For uniformly distributed data we expect that each leaf node will contain
-$n_{\textrm{node}}$ points.
-$$n_{\textrm{node}} = \frac{n}{2^{l\cdot d}}$$
+$n_{\textrm{leaf}}$ points.
+$$n_{\textrm{leaf}} = \frac{n}{2^{l\cdot d}}$$
 
 In this case $n = 10^5$, $d = 2$ and $l = 4$. So the expected number of points within a
-node is $n_{\textrm{node}} = \frac{10^5}{2^{4\cdot 2}} =390.625$.
+node is $n_{\textrm{leaf}} = \frac{10^5}{2^{4\cdot 2}} =390.625$.
 
@@
   }
-  \caption{ Distribution of node length }
+  \caption{ Distribution of leaf node length }
   \label{uniform-2d-hist}
@@
 
-The curse of high dimensionality can be deduced by the formula of $n_{\textrm{node}}$.
-The number of possible leaf nodes is $2^{l\cdot d}$ which grows exponentialy with $l$ and
-$d$, meaning that for larger $d$ the total number of points $n$ must also grow exponentialy
-to keep the ratio constant.
+Graph \ref{uniform-2d-hist} agrees with the theoreticaly expected value, although there
+is some variance. This is to be expected, increasing the number of points will reduce
+the variance then the histogram will tend more to the expected singular value.
 
 \subsection{Uniform data 10 dimensions}
-
-\begin{lstlisting}[language = Julia]
-julia> d, n = size(X)
-(10, 100000)
-
-julia> l = 1;
-
-julia> @time t = regural_bin(UInt128,
-                             X,
-                             l,
-                             1;
-                             dims=2);
-0.046737 seconds (2.48 k allocations: 30.789 MiB)
-\end{lstlisting}
-
 \subsection{MNIST data 10 dimensions}
@@
 
+% TODO Remove
+\pagebreak
+\section{Juliacon stuff, kept for reference}
 \section{Introduction}

### 2023-07-05 paper/paper.tex
chore: Paper additions PBMC MNIST

@@
   \centerline{
-    \includegraphics[width=20pc]{figures/experiments/b_vs_d.png}
+    \includegraphics[width=10pc]{figures/experiments/b_vs_d.png}
   }
@@
   \centerline{
-    \includegraphics[width=20pc]{figures/experiments/scale_n.png}
+    \includegraphics[width=10pc]{figures/experiments/scale_n.png}
   }
@@
   \centerline{
-    \includegraphics[width=20pc]{figures/experiments/scale_t.png}
+    \includegraphics[width=10pc]{figures/experiments/scale_t.png}
   }
@@
   \centerline{
-    \includegraphics[width=20pc]{figures/uniform_2d/hist.png}
+    \includegraphics[width=10pc]{figures/uniform_2d/hist.png}
   }
@@
 
-\subsection{Uniform data 10 dimensions}
-\subsection{MNIST data 10 dimensions}
-\subsection{PBMC data 10 dimensions}
-\subsection{PBMC 36 dimensions}
+\subsection{PBMC 30 dimensions}
+
+The PBMC dataset exhibits outliers, which introduces challenges in the
+normalization process. It causes the majority of data points to be clustered
+into a single node for the first three levels of the tree. As depicted in
+Figures \ref{pbmc-30-hist} (a), (b), and (c), the majority of points are
+concentrated within a maximum of two nodes at these initial levels.
+\\\\
+However, the fourth and final level of the tree highlights the impact of the
+dataset's high dimensionality. As indicated by Figure \ref{pbmc-30-hist} (d),
+it becomes apparent that most points are now assigned to nodes with a length of
+only $1$.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=10pc]{figures/experiments/pbmc_30/hist_1.png}
+    \includegraphics[width=10pc]{figures/experiments/pbmc_30/hist_2.png}
+  }
+  \centerline{
+    \includegraphics[width=10pc]{figures/experiments/pbmc_30/hist_3.png}
+    \includegraphics[width=10pc]{figures/experiments/pbmc_30/hist_4.png}
+  }
+
+  \caption{
+    30-Dimensional PBMC data distribution of leaf node length for tree level
+    (a) 1 (b) 2 (c) 3 (d) 4.
+  }
+  \label{pbmc-30-hist}
+\end{figure}
+
+\subsection{MNIST 30 dimensions}
+
+Unlike the PBMC dataset, the MNIST dataset does not exhibit significant
+outliers. As a result, the distribution between the leaf nodes in the tree
+becomes more uniform. Figure \ref{mnist-30-hist} (a) reveals that a majority of
+points are concentrated within leaf nodes that possess sufficiently large
+lengths. This favorable grouping indicates a more balanced distribution of data
+points.
+\\\\
+However, as we progress to higher levels of the tree, the impact of the
+dataset's high dimensionality becomes apparent. Consequently, the points become
+exponentially sparser, leading to the formation of singleton leaf nodes.
+This observation is clearly demonstrated in Figure \ref{mnist-30-hist} (b).
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=10pc]{figures/experiments/mnist_30/hist_1.png}
+    \includegraphics[width=10pc]{figures/experiments/mnist_30/hist_2.png}
+  }
+
+  \caption{
+    30-Dimensional MNIST data distribution of leaf node length for tree level
+    (a) 1 (b) 2
+  }
+  \label{mnist-30-hist}
+\end{figure}
+
+Figure \ref{mnist-30-hist-th} indicates that adjusting the \verb|Small|
+threshold to approximately $10000$ causes a shift in nodes from the region of
+$\geq 10000$ to the histogram's small node region, while leaving the rest of
+the nodes intact. This outcome is expected because nearly all points in the
+$\geq 10000$ region are mapped to a singleton node in the subsequent recursion.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=10pc]{figures/experiments/mnist_30/hist_3.png}
+  }
+
+  \caption{
+    30-Dimensional MNIST data distribution of leaf node length for tree level
+    2. Using Small threshold of 10000.
+  }
+  \label{mnist-30-hist-th}
+\end{figure}
 

### 2023-07-08 paper/paper.tex
chore: Paper additions

@@
 
-\section{Overview}
-To partition a cloud of points of $d$ dimensions,
-\verb|AdaptiveHierarchicalRegularBinning.jl| maps each point to a $1$ dimensional
-space filling curve, Morton or Z-order curve. The hierarchical tree is defined by the
-Morton index of each point.
+\section{The need for space partitioning algorithms}
+Space partitioning algorithms play a pivotal role in improving the runtime
+performance of computational algorithms by dividing data into smaller,
+manageable sections. These algorithms are designed to optimize the performance
+of computational algorithms by efficiently partitioning the data space in a
+manner that aligns with the specific requirements of the problem domain.
+\\\\
+In this section, we delve into the need for specific space partitioning
+algorithms and their significance in addressing the computational bottlenecks
+faced by certain algorithms. We examine three prominent application domains:
+k-nearest neighbors (kNN) search, collision detection, and the fast multipole
+method (FMM). These domains represent scenarios where the all-vs-all model,
+involving computations between every pair of data points, hampers the runtime
+performance of algorithms.
+
+\subsection{kNN}
+One common challenge encountered in computational algorithms, such as k-nearest
+neighbors (kNN) search, is the need to compute distances between each pair of
+points in the data space. This all-vs-all model becomes computationally
+expensive as the dataset size increases. However, by employing a space
+partitioning algorithm, the number of distance computations can be
+significantly reduced. Instead of calculating distances between all points,
+the algorithm subdivides the space into smaller sections, enabling the
+identification of nearest neighbors within localized partitions. This approach
+drastically improves the runtime performance of the kNN algorithm by reducing
+the computational burden associated with distance calculations.
+
+\subsection{Collision Detection}
+Collision detection algorithms also stand to benefit from specific space
+partitioning techniques. In scenarios where it is necessary to determine if a
+collision occurs between objects in a three-dimensional space, the all-vs-all
+model becomes impractical for large-scale simulations. By partitioning the
+space into smaller regions or bins, collision detection algorithms can focus
+computations solely on objects within the same partition. This significantly
+reduces the number of pairwise collision checks required and provides a
+substantial performance improvement in terms of runtime. Specific space
+partitioning algorithms enable efficient identification of potential collision
+candidates, contributing to faster and more accurate collision detection in
+simulations and virtual environments.
+
+\subsection{FMM}
+
+The fast multipole method (FMM), explored in \cite{FMM} and \cite{AMM}, is an
+important algorithm used for solving the N-body problem, which involves the
+interaction of multiple particles in a system. The traditional implementation
+of FMM suffers from a high computational complexity, as it requires pairwise
+computations between all particles in the system. However, space partitioning
+algorithms offer an opportunity to optimize the FMM by dividing the space into
+subregions or bins. By employing space partitioning techniques, the FMM
+algorithm can efficiently identify distant interactions, reducing the number of
+pairwise computations required and significantly improving the overall runtime
+performance. The incorporation of space partitioning algorithms in FMM enables
+the efficient solution of the N-body problem and opens doors for the simulation
+of large-scale physical systems.
+
+\section{Implementation}
+The partitioning of a $d$-dimensionsional cloud of points is a fundamental task
+in many computational applications. The
+\verb|AdaptiveHierarchicalRegularBinning.jl|
+algorithm offers an effective approach to address this challenge. By utilizing
+the Morton curve, a type of one-dimensional space filling curve, this algorithm
+maps each point from the $d$-dimensional space to a one-dimensional
+representation. Consequently, the hierarchical tree structure is established
+based on the Morton index assigned to each point.
+\\\\
+The \verb|AdaptiveHierarchicalRegularBinning.jl| algorithm leverages the
+inherent properties of the Morton curve to efficiently partition the cloud of
+points. The Morton curve exhibits a desirable characteristic of spatial
+locality preservation, where nearby points in the original $d$-dimensional
+space tend to have adjacent Morton indices in the one-dimensional space. This
+property ensures that points with similar spatial coordinates are grouped
+together in the resulting hierarchical tree structure.
 
@@
 
-  \begin{figure}
+  \begin{figure}[!ht]
     \centerline{
@@
 \subsection{Reordering}
-Sorting in terms of the morton indices groups the points together for efficient access
-when the tree is iterated over. We use the \verb|MSD Radixsort| algorithm to perform the
-soring.
+Sorting the points based on their Morton indices is a key operation in the
+\verb|AdaptiveHierarchicalRegularBinning.jl| algorithm, enabling efficient
+access during tree iteration. To achieve this, the algorithm employs the
+\verb|MSD Radixsort| \cite{radixsort} algorithm, or Most Significant Digit
+Radixsort, which examines the most significant bits of the Morton indices. By
+grouping points with similar prefixes together, the sorted order achieved
+through \verb|MSD Radixsort| enhances data locality and improves performance.
+This arrangement ensures that neighboring points in the original $d$-dimensional
+space are adjacent in the sorted sequence, optimizing point retrieval and
+overall algorithm efficiency.
 
@@
 with the same most significant digit (MSD) form a consecutive sub-array. The algorithm
-recurses for each such sub-array selecting the next digit in line. Selecting a large
-radix is crucial for the algorithm since a large radix lowers the number of recursions.
+recurses for each such sub-array selecting the next digit in line, leading to
+$O(n\cdot m)$ runtime complexity, where $n$ is the size of the array and $m$ is
+the number of recursions. Selecting a large radix is crucial for the algorithm
+since a large radix lowers the number of recursions.
 \\\\
@@
 Our implementation utilizes a watefall scheme to adapt to the needs of the problem at
-hand. There are $3$ different thresholds that can tune the runtime performance of the
+hand. There are three different thresholds that can tune the runtime performance of the
 spatial partitioning.
@@
 
-\subsection{Curse of high dimensionality}
-For $n=10^5$ uniformly distributed points we computed the average number of points in a
-leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
-dimensional spaces the data get exponatialy sparser.
-
-\begin{figure}[!ht]
-  \centerline{
-    \includegraphics[width=10pc]{figures/experiments/b_vs_d.png}
-  }
-  \caption{ Exponential decrease of mean number of points per leaf node }
-  \label{b_vs_d}
-\end{figure}
-
 \subsection{Runtime performance}
@@
 
-\subsection{Uniform data 2 dimensions}
+\subsection{Point per node distribution}
 
-\begin{lstlisting}[language = Julia]
-julia> d, n = size(X)
-(2, 100000)
-
-julia> l = 4;
-
-julia> @time t = regural_bin(UInt128,
-                             X,
-                             l,
-                             1;
-                             dims=2);
-0.025399 seconds (... 43.93% gc time)
-\end{lstlisting}
+\subsubsection{Uniform data 2 dimensions}
 
@@
   \centerline{
-    \includegraphics[width=10pc]{figures/uniform_2d/hist.png}
+    \includegraphics[width=10pc]{figures/experiments/uniform_2d/hist.png}
   }
@@
 
-\subsection{PBMC 30 dimensions}
+\subsubsection{Curse of high dimensionality}
+For $n=10^5$ uniformly distributed points we computed the average number of points in a
+leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
+dimensional spaces the data get exponatialy sparser.
+
+\begin{figure}[!ht]
+  \centerline{
+    \includegraphics[width=10pc]{figures/experiments/b_vs_d.png}
+  }
+  \caption{ Exponential decrease of mean number of points per leaf node }
+  \label{b_vs_d}
+\end{figure}
+
+\subsubsection{PBMC 30 dimensions}
 
@@
   \caption{
-    30-Dimensional PBMC data distribution of leaf node length for tree level
-    (a) 1 (b) 2 (c) 3 (d) 4.
+    30-Dimensional PBMC data distribution of leaf node length per leaf for tree
+    level (a) 1 (b) 2 (c) 3 (d) 4.
   }
@@
 
-\subsection{MNIST 30 dimensions}
+\subsubsection{MNIST 30 dimensions}
 
@@
 \\\\
-However, as we progress to higher levels of the tree, the impact of the
+However, as we progress to lower levels of the tree, the impact of the
 dataset's high dimensionality becomes apparent. Consequently, the points become
@@
   \caption{
-    30-Dimensional MNIST data distribution of leaf node length for tree level
-    (a) 1 (b) 2
+    30-Dimensional MNIST data distribution of leaf node length per point for
+    tree level (a) 1 (b) 2
   }
@@
   \caption{
-    30-Dimensional MNIST data distribution of leaf node length for tree level
-    2. Using Small threshold of 10000.
+    30-Dimensional MNIST data distribution of leaf node length per point for
+    tree of max level 2. Using Small threshold of 10000.
   }
@@
 \end{figure}
-
-% TODO Remove
-\pagebreak
-\section{Juliacon stuff, kept for reference}
-\section{Introduction}
-
-The \LaTeX{} document preparation system is a special version of the
-\TeX; typesetting program where-in a collection of \TeX; commands
-are added to \LaTeX{} to simplify typesetting. Importantly, it allows
-the author to concentrate on the logical structure of the document
-rather than its visual layout.\vskip 6pt
-Moreover, \LaTeX{} provides a consistent and comprehensive document
-preparation interface. There are simple-to-use commands for
-generating a table of contents, lists of figures and/or tables, and indexes.
-\LaTeX{} can automatically number list entries, equations, figures,
-tables, and footnotes, as well as articles, sections, and subsections.
-Using this numbering system, bibliographic citations, page
-references, and cross references to any other numbered entity (e.g.
-article, section, equation, figure, list entry, etc.) become quite simple
-and straightforward. The use of \LaTeX{} document classes allows
-a simple change of class to transform the appearance of your document.\vskip 6pt
-\LaTeX{} is a powerful tool for managing long and complex documents.
-In particular, partial processing enables long documents to
-be produced article by article without losing sequential information.
-The use of document classes allows a simple change of style
-(or style option) to transform the appearance of your document.
-
-\section{The JuliaCon Article Class}
-\label{sec:documentclass}
-%
-The juliacon class file preserves the standard LATEX{} interface such
-that any document that can be produced using the standard LATEX{}
-article class can also be produced with the class file.\vskip 6pt
-It is likely that the make up will change after file submission. For
-this reason, we ask you to ignore details such as slightly long lines,
-page stretching, or figures falling out of synchronization, as these
-details can be dealt with at a later stage.\vskip 6pt
-Use should be made of symbolic references (\verb|\ref| ) in order to
-protect against late changes of order, etc.
-
-\section{USING THE JuliaCon Article CLASS FILE}
-
-If the file \verb|juliacon.cls|  is not already in the appropriate system directory
-for \LaTeX{} files, either arrange for it to be put there or copy
-it to your working directory. The \verb|juliacon|  document class is implemented
-as a complete class, not a document style option. In order to
-use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
-\verb|\documentclass|  command at the beginning of your document:
-\vskip 6pt
-\begin{centering}
-    \verb|\documentclass{article}|  \end{centering}
-\vskip 6pt
-replace by
-\vskip 6pt
- \verb|\documentclass{juliacon}|  \vskip 6pt
-In general, the following standard document \verb|style|  options should
-{ \itshape not} be used with the {\footnotesize \itshape article} class file:
-\begin{enumerate}
-\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
-\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
-\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
-\end{enumerate}
-
-\section{Additional Document Style Options}
-\label{sec:additional_doc}
-%
-The following additional style option is available with the \verb|juliacon|  class file:
-\vskip 6pt
-Please place any additional command definitions at the very start of
-the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
-\verb|\def|  and \verb|\newcommand|   commands that define macros for
-technical expressions should be placed here. Other author-defined
-macros should be kept to a minimum.
-\vskip 6pt
-Commands that differ from the standard \LaTeX{} interface, or that
-are provided in addition to the standard interface, are explained in
-this guide. This guide is not a substitute for the \LaTeX{} manual itself.
-Authors planning to submit their papers in \LaTeX{} are advised to use
-\verb|juliacon.cls|  as early as possible in the creation of their files.
-
-%
-%
-%
-%
-\begin{table*}[t]
-\tabcolsep22pt
-\tbl{If necessary, the tables can be extended both columns.}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\label{tab:symbols}
-\begin{tabnote}
-This is an example of table footnote.
-\end{tabnote}
-\end{table*}
-% \begin{figure*}[t]
-% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
-% \caption{If necessary, the images can be extended both columns.}
-%   \label{fig:sample_image}
-% \end{figure*}
-
-\section{Additional features}
-\label{sec:additional_faci}
-In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
-In general, once you have used the additional \verb|juliacon.cls| facilities
-in your document, do not process it with a standard \LaTeX{} class
-file.
-
-\subsection{Titles, Author's Name, and Affiliation}
-\label{subsub:title_auth}
-The title of the article, author's name, and affiliation are used at the
-beginning of the article (for the main title). These can be produced
-using the following code:
-
-\begin{verbatim}
-\title{ This is an example of article title} }
-\author{
-   \large 1st Author \\[-3pt]
-   \normalsize 1st author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	1st author's email address \\[-3pt]
-  \and
-   \large 2nd Author \\[-3pt]
-   \normalsize 2nd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	2nd author's email address \\[-3pt]
-\and
-   \large 3rd Author \\[-3pt]
-   \normalsize 3rd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	3rd author's email address \\[-3pt]
-}
-\maketitle
-\end{verbatim}
-
-\subsection{Writing Julia code}
-
-A special environment is already defined for Julia code,
-built on top of \textit{listings} and \textit{jlcode}.
-
-\begin{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-\end{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-
-\subsection{Abstracts, Key words, term etc...}
-\label{subsub:abs_key_etc}
-
-At the beginning of your article, the title should be generated
-in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
-\verb|\terms|,
-\verb|\keywords|  commands respectively. The abstract should be enclosed
-within an abstract environment, All these environment
-can be produced using the following code:
-\begin{verbatim}
-\terms{Experimentation, Human Factors}
-
-\keywords{Face animation, image-based modelling...}
-
-\begin{abstract}
-In this paper, we propose a new method for the
-systematic determination of the model's base of
-time varying delay system. This method based on
-the construction of the classification data related
-to the considered system. The number, the orders,
-the time delay and the parameters of the local
-models are generated automatically without any
-knowledge about the full operating range of the
-process. The parametric identification of the local
-models is realized by a new recursive algorithm for
-on line identification of systems with unknown time
-delay. The proposed algorithm allows simultaneous
-estimation of time delay and parameters of
-discrete-time systems. The effectiveness of
-the new method has been illustrated through
-simulation.
-\end{abstract}
-
-\end{verbatim}
-
-\section{Some guidelines}
-\label{sec:some_guide}
-The following notes may help you achieve the best effects with the
-\verb|juliacon|  class file.
-
-\subsection{Sections}
-\label{subsub:sections}
-\LaTeXe{}  provides four levels of section headings and they are all
-defined in the \verb|juliacon|  class file:
-\begin{itemize}
-\item \verb|\section|
-\item \verb|\subsection|
-\item \verb|\subsubsection|
-\item \verb|\paragraph|
-\end{itemize}
-Section headings are automatically converted to allcaps style.
-\subsection{Lists}
-\label{sec:lists}
-%
-The \verb|juliacon|   class file provides unnumbered lists using the
-\verb|unnumlist|  environment for example,
-
-\begin{unnumlist}
-\item First unnumbered item which has no label and is indented from the
-left margin.
-\item Second unnumbered item.
-\item Third unnumbered item.
-\end{unnumlist}
-The unnumbered list which has no label and is indented from the
-left margin. was produced by:
-\begin{verbatim}
-\begin{unnumlist}
-\item First unnumbered item...
-\item Second unnumbered item...
-\item Third unnumbered item...
-\end{unnumlist}
-\end{verbatim}
-
-The \verb|juliacon|  class file also provides hyphen list using the
-\verb|itemize|  environment for example,
-\begin{itemize}
-\item First unnumbered bulleted item which has no label and is indented
-from the left margin.
-\item Second unnumbered bulleted item.
-\item Third unnumbered bulleted item which has no label and is indented
-from the left margin.
-\end{itemize}
-was produced by:
-\begin{verbatim}
-\begin{itemize}
-\item First item...
-\item Second item...
-\item Third item...
-\end{itemize}
-\end{verbatim}
-
-Numbered list is also provided in acmtog class file using the
-enumerate environment for example,
-\begin{enumerate}
-\item The attenuated and diluted stellar radiation.
-\item Scattered radiation, and
-\item Reradiation from other grains.
-\end{enumerate}
-
-was produced by:
-\begin{verbatim}
-\begin{enumerate}
-\item The attenuated...
-\item Scattered radiation, and...
-\item Reradiation from other grains...
-\end{enumerate}
-\end{verbatim}
-\subsection{Illustrations (or figures)}
-\label{subsub:sec_Illus}
-The \verb|juliacon|  class file will cope with most of the positioning of
-your illustrations and you should not normally use the optional positional
-qualifiers on the \verb|figure|  environment that would override
-these decisions.
-\vskip 6pt
-
-%
-\begin{figure}[t]
-\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
-\caption{This is example of the image in a column.}
-	\label{fig:sample_figure}
-\end{figure}
-
-The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
-organization \footnote{https://github.com/JuliaGraphs}.
-
-Figure captions should be \emph{below} the figure itself, therefore the
-\verb|\caption|  command should appear after the figure or space left for
-an illustration. For example, Figure 1 is produced using the following
-commands:
-
-\begin{verbatim}
-\begin{figure}
-\centerline{\includegraphics[width=20pc]{Graphics.eps}}
-\caption{An example of the testing process for a
-binary tree. The globa null hypothesis is tested
-first at level $\alpha$ (a), and the level of
-individual variables is reached last (d). Note
-that individual hypotheses can be tested at
-level $\alpha/4$ and not $\alpha/8$ as one might
-expect at first.}
-\label{sample-figure_2}
-\end{figure}
-\end{verbatim}
-Figures can be resized using first and second argument of
-\verb|\includegraphics|   command. First argument is used for modifying
-figure height and the second argument is used for modifying
-figure width respectively.
-\vskip 6pt
-Cross-referencing of figures, tables, and numbered, displayed
-equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
-For example, in referencing Figure 1 above, we used
-\verb|Figure~\ref{sample-figure}|
-
- \subsection{Tables}
-\label{subsub:sec_Tab}
-The \verb|juliacon|   class file will cope with most of the positioning of
-your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
-decisions. Table captions should be at the top.
-\begin{verbatim}
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Train70 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test70 & Testing Data I & & 105\\\hline
-Test30 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\end{verbatim}
-
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\subsection{Landscaping Pages}
-\label{subsub:landscaping_pages}
-If a table is too wide to fit the standard measure, it may be turned,
-with its caption, to 90 degrees. Landscape tables cannot be produced
-directly using the \verb|juliacon|   class file because \TeX{} itself cannot
-turn the page, and not all device drivers provide such a facility.
-The following procedure can be used to produce such pages.
-\vskip 6pt
-Use the package \verb|rotating|   in your document and change the coding
-from
-\begin{verbatim}
-\begin{table}...\end{table}
-to
-\begin{sidewaystable}...\end{sidewaystable}
-and for figures
-\begin{figure}...\end{figure}
-to
-\begin{sidewaysfigure}...\end{sidewaysfigure}
-\end{verbatim}
-
-environments in your document to turn your table on the appropriate
-page of your document. For instance, the following code prints
-a page with the running head, a message half way down and the
-table number towards the bottom.
-\begin{verbatim}
-\begin{sidewaystable}
-\tbl{Landscape table caption to go here.}{...}
-\label{landtab}
-\end{sidewaystable}
-\end{verbatim}
-
-\subsection{Double Column Figure and Tables}
-\label{subsub:double_fig_tab}
-For generating the output of figures and tables in double column
-we can use the following coding:
-
-\begin{enumerate}
-\item For Figures:
-\begin{verbatim}
-\begin{figure*}...\end{figure*}
-\end{verbatim}
-\item For landscape figures:
-\begin{verbatim}
-\begin{sidewaysfigure*}...\end{sidewaysfigure*}
-\end{verbatim}
-\item For Tables:
-\begin{verbatim}
-\begin{table*}...\end{table*}
-\end{verbatim}
-\item For landscape tables:
-\begin{verbatim}
-\begin{sidewaystable*}...\end{sidewaystable*}
-\end{verbatim}
-\end{enumerate}
-
-\subsection{Typesetting Mathematics}
-\label{subsub:type_math}
-The \verb|juliacon| class file will set displayed mathematics with center to
-the column width, provided that you use the \LaTeXe{} standard of
-open and closed square brackets as delimiters.
-The equation
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-
-was typeset using the acmtog class file with the commands
-
-\begin{verbatim}
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-\end{verbatim}
-
-For display equations, cross-referencing is encouraged. For example,
-\begin{verbatim}
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-\end{verbatim}
-The following output is generated with the above coding:
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-
-\subsection{Enunciations}
-\label{subsub:enunciation}
-The \verb|juliacon|   class file generates the enunciations with the help of
-the following commands:
-\begin{verbatim}
-\begin{theorem}...\end{theorem}
-\begin{strategy}...\end{strategy}
-\begin{property}...\end{property}
-\begin{proposition}...\end{proposition}
-\begin{lemma}...\end{lemma}
-\begin{example}...\end{example}
-\begin{proof}...\end{proof}
-\begin{definition}...\end{definition}
-\begin{algorithm}...\end{algorithm}
-\begin{remark}...\end{remark}
-\end{verbatim}
-The above-mentioned coding can also include optional arguments
-such as
-\begin{verbatim}
-\begin{theorem}[...]. Example for theorem:
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven ... created equal.
-\end{theorem}
-\end{verbatim}
-
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven years ago our fathers brought forth,
-upon this continent, a new nation, conceived in Liberty,
- and dedicated to the proposition that all men are
-created equal.
-\end{theorem}
-
-\subsection{Extract}
-\label{subsub:extract}
-Extract environment should be coded within
-\begin{verbatim}
-\begin{extract}..\end{extract}
-\end{verbatim}
-
-\subsection{Balancing column at last page}
-\label{subsub:Balance}
-For balancing the both column length at last page use :
-\begin{verbatim}
-\vadjust{\vfill\pagebreak}
-\end{verbatim}
-
-%\vadjust{\vfill\pagebreak}
-
-at appropriate place in your \TeX{} file or in bibliography file.
-
-\section{Handling references}
-\label{subsub:references}
-References are most easily (and correctly) generated using the
-BIBTEX, which is easily invoked via
-\begin{verbatim}
-\bibliographystyle{juliacon}
-\bibliography{ref}
-\end{verbatim}
-When submitting the document source (.tex) file to external
-parties, the ref.bib file should be sent with it.
-\cite{bezanson2017julia}
 

### 2023-07-08 paper/paper.tex
chore: Additions to paper

@@
 
-\section{The need for space partitioning algorithms}
+\section{Motivation}
 Space partitioning algorithms play a pivotal role in improving the runtime
@@
 \\\\
-In this section, we delve into the need for specific space partitioning
+In this section, we delve into the need for space partitioning
 algorithms and their significance in addressing the computational bottlenecks
@@
 \subsection{Collision Detection}
-Collision detection algorithms also stand to benefit from specific space
+Collision detection algorithms also stand to benefit from space
 partitioning techniques. In scenarios where it is necessary to determine if a
-collision occurs between objects in a three-dimensional space, the all-vs-all
+collision occurs between objects in a $d$-dimensional space, the all-vs-all
 model becomes impractical for large-scale simulations. By partitioning the
